---
title: 'Bios 6301: Assignment 2'
output: pdf_document
author: "Yudong Cao"
---

*(informally) Due Tuesday, 20 September, 1:00 PM*

50 points total.

This assignment won't be submitted until we've covered Rmarkdown.
Create R chunks for each question and insert your R code appropriately.
Check your output by using the `Knit PDF` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
    ```{r}
    setwd("~/Desktop/FA 16/BIOS 6301/hw")
    cancer.df<-read.csv("cancer.csv")
    cancer.df<-data.frame(cancer.df)
    attach(cancer.df)
    ```

    2. Determine the number of rows and columns in the data frame. (2)
    ```{r}
    nrow(cancer.df)
    ncol(cancer.df)
    ```

    3. Extract the names of the columns in `cancer.df`. (2)
    ```{r}
    names(cancer.df)
    ```

    4. Report the value of the 3000th row in column 6. (2)
    ```{r}
    cancer.df[3000,6]
    ```

    5. Report the contents of the 172nd row. (2)
    ```{r}
    cancer.df[172,]
    ```

    6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)
    ```{r}
    cancer.df[,'incrate']<-incidence/100000
    ```

    7. How many subgroups (rows) have a zero incidence rate? (2)
    ```{r}
    sum(cancer.df['incrate']==0)
    #23191 rows have zero incrate
    ```

    8. Find the subgroup with the highest incidence rate.(3)
    ```{r}
    which(cancer.df['incrate']==max(cancer.df['incrate']))
    #the 21387th subgroup has highest incrate
    ```

2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)
    ```{r}
    x <- c("5","12","7")
    max(x)
    # no error; for characters, the first 'number' is compared to get max(x)=7
    sort(x)
    # no error; for characters, the first 'number' is primarily sorted "12" "5" "7"
    
    #sum(x)
    # error; cannot sum up elements of character strings
    ```

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)
    ```{r}
    y <- c("5",7,12)
    # define y to be a vector with the character string "5", "7" and "12"
    
    #y[2] + y[3]
    # error because R cannot add two elements of character strings
    ```
    
    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)
    ```{r}
    z <- data.frame(z1="5",z2=7,z3=12)
    # define z to be a data frame with a row of numeric elements 5, 7 and 12
    z[1,2] + z[1,3]
    # add up the 2nd and 3rd element of z, 7+12=19
    ```
    
3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
    ```{r}
    c(1:8,7:1)
    ```
    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
    ```{r}
    rep(1:5,time=1:5)
    ```

    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
    ```{r}
    1-diag(3)
    ```

    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
    ```{r}
    y=c(1:4)
    x=matrix(c(y,y^2,y^3,y^4,y^5),nrow=5,byrow=T)
    x
    ```

4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)
```{r}
h<-function(x,n) {
  h<-1
  for (i in 1:n) {
    h<-h+x^i
  }
  return(h)
}
```

    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
```{r}
s<-0
for (i in 1:999) {
  if (i%%3==0 | i%%5==0) {
    s<-s+i
  }
}
s
```
        
        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
```{r}
t<-0
for (j in 1:999999) {
  if (j%%4==0 | j%%7==0) {
    t<-t+j
  }
}
t
```

    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
```{r}
x<-c(1,2)
for (i in 1:100) {
  x[i+2]<-x[i]+x[i+1]
}
y<-0
for (j in 1:45) {
  if (x[j]%%2==0) {
    y<-y+x[j]
  }
}
y
```

Some problems taken or inspired by projecteuler.

[euler1]:https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
